Copulas and their potential for ecology

### 2019-09-20 Paper.Rmd
edits done as suggested by DCR: see my response

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*; copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the section on Background below. Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
respectively. See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the section on Background below, 
after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals \cite{Genest2007}. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables; 
they provide information solely on the dependence between those variables.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7in]{./MasterFigure.pdf}
\caption{Summary and guide for analyses presented in the text. Middle boxes
correspond to ecological datasets for which Q1 was examined (in the 
third through
fifth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
correspond to potential causes of non-normal (non-N) copula structure that
were examined (in the sixth and seventh sections of the text). Lower boxes 
correspond to potential consequences we examined of non-N copula structure for
ecological understanding and for applications (in the eighth and ninth
sections of the text). Arrow labels (A-D for analyses pertaining 
to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
link to locations in the text.\label{MasterFigure}}
\end{center}
\end{figure}

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family (henceforth abbreviated N), which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of dependence
between the variables. An N copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate N copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails (see
below for a more precise defintion of tail associations). 
The Clayton copula family (henceforth abbreviated C),
of which we also already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to N copulas, C copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
Clayton family (here abbreviated SC) is the symmetric opposite of the C family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

@@

Having introduced tail association conceptually and referred to it
in examples, we now need a more precise definition of a measure of tail association.
We use the term *tail association* to describe the general idea of the strength of
associations between two variables in the tails of their distributions. One of the
ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
dependence of $X$ and $Y$ is defined as 
$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals 
$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
which shows upper-tail dependence is defined symmetrically in the two random variables.
Lower tail dependence is defined analogously, as 
$\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
Tail dependence is a property of the random variable $(U,V)$, so depends only
on the copula and not on the marginals combined with it. It can be shown 
(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the N copula,
$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a C copula with parameter $p>0$,
and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an SC copula with parameter $p>0$.
In the section on methods for Q1 we will introduce
ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
introduce nonparametric measures of tail association that apply directly to data.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; $p$ is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition). The 
parameter range for the family is $p \in [-1,1]$, and lower- and upper-tail
dependence are always $0$.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
copulas. S is Spearman correlation; $p$ is the value of the parameter for the C family (it 
is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, 
respectively (see Background section for definition). The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
dependence is $0$.\label{figped_C}}
\end{center}
\end{figure}

@@

We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the N
copula, making possible comparisons of the degree to which the N copula versus
other copula families were good descriptions of data.

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data \cite{Genest2007}. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of locations, and 
compared AIC weights and tail-dependence of 
minimum-AIC copulas with that of an N copula. We also counted the number of location pairs 
for which model-averaged lower-tail dependence was 
greater than model-averaged upper-tail dependence 
or vice-versa, and reported the percentage of non-independent 
location pairs showing stronger lower-tail dependence. We carried out
pairwise bivariate analyses for simplicity and because they were
sufficient to answer our research questions. But see the Discussion
for a few words on multivariate copulas.

@@

To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
project is archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--***SHYAMOLINA: Please do the following to help me revise this section, so far only covering changes in the presentation of the bivariate results. Please let me know if any of this is unclear or you want to discuss anything.
1) Replace the pithy abbreviations for the different copulas (N, C, SC, G, etc.) with their full names throughout the main text and sup mat, including in figure and table captions. This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.
2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.
3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
7) Combine tables S6, S8, S10 into one table, and fix it so
whenever we currently cite one of those, change it so we cite the combined table instead.
8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.
-->

@@

Across essentially all our datasets, copula structures were typically non-N and 
commonly showed asymmetric tail 
dependence, answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we summarize for all multivariate datasets.

<!--bivariate example-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{SM-tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
N copula had AIC weight essentially 0, answering the 
first part of Q1 for these data. 
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what an N copula would give, namely, 0, answering the second part 
Q1 for these data. The values also differed substantially 
from each other, answering the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

Although these results convincingly show non-N copula structure, 
we caution that the model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`). However, nonparametric statistics verified
both that tail dependence was distinct from that of an N copula, and that it 
was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). Our 
affirmative, empirically based answer to Q1 for the soil C and N data is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-N copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The N copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first
part of Q1. 
Lower- and upper-tail dependence statistics often differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
answering the second and third parts of Q1. 
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though weakly. 
Although these results convincingly show non-N copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
Our 
affirmative, empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of those
datasets, in the middle row of boxes.
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000: the best-fitting copula was not N, and there was slightly stronger lower- 
than upper-tail dependence. Detailed results for each dataset are in 
Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.

>>>>

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*; copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the section on Background below. Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
respectively. See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the section on Background below, 
after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables; 
they provide information solely on the dependence between those variables.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7in]{./MasterFigure.pdf}
\caption{Summary and guide for analyses presented in the text. Middle boxes
correspond to ecological datasets for which Q1 was examined (in the 
third through
fifth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
correspond to potential causes of non-normal (non-N) copula structure that
were examined (in the sixth and seventh sections of the text). Lower boxes 
correspond to potential consequences we examined of non-Normalcopula structure for
ecological understanding and for applications (in the eighth and ninth
sections of the text). Arrow labels (A-D for analyses pertaining 
to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
link to locations in the text.\label{MasterFigure}}
\end{center}
\end{figure}

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family (henceforth abbreviated N), which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of dependence
between the variables. An Normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate Normalcopulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails (see
below for a more precise defintion of tail associations). 
The Clayton copula family (henceforth abbreviated C),
of which we also already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to Normalcopulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
Clayton family (here abbreviated SC) is the symmetric opposite of the Claytonfamily, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

@@

Having introduced tail association conceptually and referred to it
in examples, we now need a more precise definition of a measure of tail association.
We use the term *tail association* to describe the general idea of the strength of
associations between two variables in the tails of their distributions. One of the
ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
dependence of $X$ and $Y$ is defined as 
$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals 
$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
which shows upper-tail dependence is defined symmetrically in the two random variables.
Lower tail dependence is defined analogously, as 
$\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
Tail dependence is a property of the random variable $(U,V)$, so depends only
on the copula and not on the marginals combined with it. It can be shown 
(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the Normalcopula,
$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an survival Clayton  copula with parameter $p>0$.
In the section on methods for Q1 we will introduce
ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
introduce nonparametric measures of tail association that apply directly to data.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example Normalcopulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Normalcopulas. 
S is Spearman correlation; $p$ is the value of the parameter for the Normal
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition). The 
parameter range for the family is $p \in [-1,1]$, and lower- and upper-tail
dependence are always $0$.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example Clayton  copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Clayton  
copulas. S is Spearman correlation; $p$ is the value of the parameter for the Clayton  family (it 
is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, 
respectively (see Background section for definition). The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
dependence is $0$.\label{figped_C}}
\end{center}
\end{figure}

@@

We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the N
copula, making possible comparisons of the degree to which the Normalcopula versus
other copula families were good descriptions of data.

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data [@Genest2007]. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of locations, and 
compared AIC weights and tail-dependence of 
minimum-AIC copulas with that of an Normalcopula. We also counted the number of location pairs 
for which model-averaged lower-tail dependence was 
greater than model-averaged upper-tail dependence 
or vice-versa, and reported the percentage of non-independent 
location pairs showing stronger lower-tail dependence. We carried out
pairwise bivariate analyses for simplicity and because they were
sufficient to answer our research questions. But see the Discussion
for a few words on multivariate copulas.

@@

To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--***SHYAMOLINA: Please do the following to help me revise this section, so far only covering changes in the presentation of the bivariate results. Please let me know if any of this is unclear or you want to discuss anything.
1) Replace the pithy abbreviations for the different copulas (N, C, SC, G, etc.) with their full names throughout the main text and sup mat, including in figure and table captions. This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.

>>>> Done. I don't know whether it should be called as a Normal copula or an Normal copula? Now, we use "an".

2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.

>>>> Added, sorting based on min AIC.

3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
7) Combine tables S6, S8, S10 into one table, and fix it so
whenever we currently cite one of those, change it so we cite the combined table instead.
8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.

>>>> All done.
-->

@@

Across essentially all our datasets, copula structures were typically non-Normaland 
commonly showed asymmetric tail 
dependence, answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we summarize for all multivariate datasets.

<!--bivariate example-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
Normal copula had AIC weight essentially 0, answering the 
first part of Q1 for these data. 
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what an Normal copula would give, namely, 0, answering the second part 
Q1 for these data. The values also differed substantially 
from each other, answering the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

Although these results convincingly show non-Normal copula structure, 
we caution that the model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`). However, nonparametric statistics verified
both that tail dependence was distinct from that of an Normal copula, and that it 
was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
Fig. \ref{fig_soilCN_nonparam}, Table \ref{tab_soilCN_asym}). Our 
affirmative, empirically based answer to Q1 for the soil C and N data is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-Normal copula 
always emerged as the lowest-AIC (highest-AIC-weight) copula 
(Table \ref{tab_bivar_summary}, row 2). 
The Normal copula was always a poor fit compared to the highest-AIC-weight copula 
(Table \ref{tab_bivar_summary}, rows 3-4), answering the first
part of Q1. 
Lower- and upper-tail dependence statistics often differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
answering the second and third parts of Q1. 
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though weakly. 
Although these results convincingly show non-Normal copula structure, the 
model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula was a poor fit in some cases 
(Table \ref{tab_bivar_summary}, rows 5-6).
But our nonparametric statistics confirmed the same conclusions 
(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
Our 
affirmative, empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of those
datasets, in the middle row of boxes.
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000: the best-fitting copula was not N, and there was slightly stronger lower- 
than upper-tail dependence. Detailed results for each dataset are in 
Tables \ref{tab_soilCNfit} - \ref{tab_soilCN_asym},  \ref{SM-tab_bmr_birds_fit} - \ref{SM-tab_3biv_npa_summary} and 
Figs. \ref{fig_soilCN_nonparam},  \ref{SM-fig_bmr_birds_nonparam} - \ref{SM-fig_cc2000_nonparam}.
